---
title: Downloading Twitter Data 
subtitle: Day 4
author: "Olga Chyzh [www.olgachyzh.com]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
exclude: true
editor_options: 
  chunk_output_type: console
---

```{r  setup, message=FALSE, warning=FALSE, include=FALSE}
options(
  htmltools.dir.version = FALSE, # for blogdown
  width = 80,
  tibble.width = 80
)

knitr::opts_chunk$set(
  fig.align = "center",  warning=FALSE, message=FALSE
)

```

## Materials

Download all materials from [Github](https://github.com/ochyzh/Websraping-Workshop)

Workshop 1 video posted [here](https://play.library.utoronto.ca/0de7f9b4b6295aafbee68a61e198c845)

Workshop 2 video posted [here](https://play.library.utoronto.ca/cc3ebb18816bfdcf8be7686bf060385b)

Workshop 3 video posted [here](https://play.library.utoronto.ca/5646bdd009c7b8f905855e8e8b272e49)

Workshop 4 video posted [here](https://play.library.utoronto.ca/3164d61a493bb8c44901d40aff512a3e)

---

## Outline

- Good data management from start to finish of the project

- Object classes: `numeric`/`integer`, `character`, `date`/`time`, `factor`.

- Most frequently used tools: `select`, `arrange`, `filter`, `mutate`, `summarise`, `join_by`

- Example: [Global Terrorism Dataset](https://www.start.umd.edu/gtd/access/). 

---

## Good Data Management Principles

- *Never* alter the raw data in any way. No sorting, fixing typos<sup>1</sup> 

    - Save all raw data files in a subfolder named `raw` and all the transformed data files in a subfolder named `data`.
    
- Save all data management in a script, so you can check/edit it as needed.

- A readable script is a short script

    - Use the pipe operators to chain functions
    
    - Loop as much as possible

- Have separate scripts for data cleaning and analysis.  

- Can you think of other principles?

.footnote[
<sup>1</sup> I do manually change Excel files to .csv format though.] 

---

## Required Packages

Install all tidyverse packages:
```{r gettv, echo = T, eval = T}
library(tidyverse)
library(magrittr)
library(lubridate)
library(ggplot2)
library(stringr)
library(data.table)
library(openxlsx)

```

---

## Get the Data

Download the full [Global Terrorism Dataset](https://www.start.umd.edu/gtd/access/) (need to provide very basic info, like your name) or a smaller version of incidents since 2010 [here](https://raw.githubusercontent.com/ochyzh/Websraping-Workshop/master/GTD_small.csv)

The full version is over 160 MB once converted to .csv. Opening it using `read.csv` will take a while. Try `data.table::fread` to read it faster:

```{r}
fulldata<-fread("./raw/globalterrorismdb_0919dist.csv", header=TRUE) 
smalldata<-fread("https://raw.githubusercontent.com/ochyzh/Websraping-Workshop/master/GTD_small.csv", header=TRUE)
```


---

## `select`, `filter` and `arrange`

- `dplyr::select` allows you to select specific columns of the data

     - E. g., use `select` to reduce the number of variables in the data

```{r}
fulldata<- select(fulldata, iyear, imonth, iday, country_txt, attacktype1_txt, latitude, longitude)
smalldata <-select(smalldata, iyear, imonth, iday, country_txt, attacktype1_txt, latitude, longitude)
```

- `dplyr::filter` allows you to select by rows

```{r, eval=F}
filter(fulldata, iyear==2018)
filter(smalldata, iyear==2018)
```

- `dplyr::arrange` allows you to sort the data

```{r}
fulldata<-arrange(fulldata, iyear, imonth, iday)
smalldata<-arrange(smalldata, iyear, imonth, iday)
```


---

## Check How Variables Are Stored

`str` is for structure

```{r, eval=F}
str(fulldata)
```

    - `int` is for integer

    - `chr` is for character

    - `num` is for numeric

- Fix any class inconsistencies, e. g. numbers stored as characters or factors

---

## Example: Convert Year/Month/Day to Date Class

To convert to `date` using `lubridate`, need:

$\color{green}\checkmark$ Year in the form XXXX 

$\color{red}\times$ Month in the form XX  

$\color{red}\times$ Day in the form XX 

$\color{red}\times$ Some months and days are coded as 0

$\color{red}\times$ Date saved in the format that combines year, month, and day, e.g., YYYY-MM-DD 

---
## Plan of Attack

[1.] Use `filter` to drop the observations with unknown day or month (can you think of other ways to treat these?)

```{r}
fulldata<- filter(fulldata, imonth!=0, iday !=0)


```

[2.] Pad month and day with leading zeroes, using `str_pad` from the `stringr` package

```{r}
fulldata$month<-str_pad(fulldata$imonth,2,pad="0")
fulldata$day<-str_pad(fulldata$iday,2,pad="0")

```

[3.] Use `paste0` to combine year, month, and date

```{r}
fulldata$date<-paste0(fulldata$iyear,fulldata$month,fulldata$day)
```

[4.] Convert variable `date` to class "date"

```{r}
fulldata$date<-ymd(fulldata$date)
```

---

## `dplyr::mutate`

- In the previous couple slides, we recoded variables in our data using `base` R.

- `mutate` is a function to recode within the `dplyr` package.

- Does the same thing, but works well with other `dplyr` functions when chained into a pipe

- Makes for shorter/clearer code chunks

```{r}
fulldata<-fulldata %>% 
  filter(imonth!=0, iday !=0) %>%
  mutate(month=str_pad(imonth,2,pad="0"), 
         day=str_pad(iday,2,pad="0"),
         date<-paste0(iyear,month,day),
         date<-ymd(date))

smalldata<-smalldata %>% 
  filter(imonth!=0, iday !=0) %>%
  mutate(month=str_pad(imonth,2,pad="0"), 
         day=str_pad(iday,2,pad="0"),
         date<-paste0(iyear,month,day),
         date<-ymd(date))
```
---
## Joining Multiple Datasets

We often want to add additional variables into our data.

For example, if we wanted to make a choropleth map of the world that colors countries based on the number of terrorist attacks in 2018, we would need to merge our data with a the data on geographical borders.

Library `mapdata` contains data on geographical borders that would allow us to make a map.

```{r}
library(mapdata)
world<-map_data("world")
head(world)
```


---
## Make the Map of the World

```{r}
ggplot(data=filter(world, region!="Antarctica"), aes(x=long, y=lat, group=group))+geom_polygon(color="black", fill="white") +
  scale_y_continuous("")
```

---

## Joining Two Datasets

- In order to join two datasets, we must identify the common variable, i.e. the "joining" variable.

```{r}
head(smalldata)
head(world)
```

- Variable `country_txt` identifies observations in teh `smalldata` dataset, while the same information in the `world` dataset is contained in the variable `region`.

- The joining variable must be formatted the same way in both datasets, e.g., capitalization, accents, apostrophes, spelling.

---
## Prepare the Terrorism Data

- The unit of observation is attack, but we want to shade countries proportional to the number of attacks.

- We need to change the level of aggregation from *attack* to *country*.

- We can use `summarise` with *country* as the grouping variable

```{r}
smalldata<- smalldata %>% 
  group_by(country_txt) %>%
  summarise(num_attks=length(country_txt), .groups="keep") %>%
  arrange(desc(num_attks))
```


---

## Check Before Joining

- Before we join these two datasets, we will do a check using `anti_join`.

- `anti_join` will show us all the countries that are not matching up.

```{r}
nomatch1<- smalldata %>% anti_join(world, by=c("country_txt"="region"))
unique(nomatch1$country_txt)

nomatch2<- world %>% anti_join(smalldata, by=c("region"="country_txt"))
unique(nomatch2$region)
```


---

## Fix Spelling of Country Names

```{r}
smalldata<-smalldata %>% mutate(region=country_txt, region=str_replace(region,"United States","USA"),
                     region=str_replace(region,"United Kingdom","UK"),
                     region=str_replace(region,"Bosnia-Herzegovina","Bosnia and Herzegovina"),
                     region=str_replace(region,"West Bank and Gaza Strip","Palestine"),
                     region=str_replace(region,"Slovak Republic","Slovakia"),
                     region=str_replace(region,"Trinidad and Tobago","Trinidad"),
                     region=str_squish(region)
                     )

```
---

## Merge

```{r}
world %>% filter(region!="Antarctica") %>% left_join(smalldata, by="region") %>%
  ggplot(aes(x=long, y=lat, group=group, fill=log(num_attks)))+geom_polygon(color="black") +
  scale_y_continuous("", breaks=NULL)+scale_x_continuous("", breaks=NULL)+ scale_fill_gradient2("Attks",na.value="white")+theme_bw()
```



---
## More resources:

    - [Karl Browman's blog](https://kbroman.org/dataorg/)
    
    - [Hadley Wickam's website](http://hadley.nz/)
    
    - [Heike Hoffman's class website](https://stat579-at-isu.github.io/)
    

