---
title: Downloading Twitter Data 
subtitle: Day 4
author: "Olga Chyzh [www.olgachyzh.com]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
exclude: true
editor_options: 
  chunk_output_type: console
---

```{r  setup, message=FALSE, warning=FALSE, include=FALSE}
options(
  htmltools.dir.version = FALSE, # for blogdown
  width = 80,
  tibble.width = 80
)

knitr::opts_chunk$set(
  fig.align = "center",  warning=FALSE, message=FALSE
)

```



## Outline

    
- Make a Twitter App

- Use `rtweet` to download Twitter data

Download all materials from [Github](https://github.com/ochyzh/Websraping-Workshop)

Workshop 1 video posted [here](https://play.library.utoronto.ca/0de7f9b4b6295aafbee68a61e198c845)

Workshop 2 video posted [here](https://play.library.utoronto.ca/cc3ebb18816bfdcf8be7686bf060385b)

Workshop 3 video posted [here](https://play.library.utoronto.ca/5646bdd009c7b8f905855e8e8b272e49)


---
## Required Packages

Install all tidyverse packages:
```{r gettv, echo = T, eval = T}
# check if you already have it
library(tidyverse)
library(magrittr)
library(rtweet)
library(lubridate)
library(ggplot2)
library(stringr)
library(igraph)

# if not:
#install.packages("rtweet")
#library(rtweet) # only calls the "core" of tidyverse

#Or install the development version:
#devtools::install_github("mkearney/rtweet")


```


---

## Some Background

- Many websites use APIs (application protocol interfaces) to manage and organize their data.

- These are often public (or allow authorized users), so we can use them to obtain data.

- Twitter, Facebook, Reddit, the World Bank website, the US Census Website

- App-specific R packages to make it easy: `rtweet`, `Rfacebook` `RedditExtractoR` 

---
## Twitter APIs

- Connecting to Twitter's API requires access tokens: **api key**, **api secret key**, **access token**, **access token secret**

- Need a Twitter Developer Account

- If don't have one, follow the instructions outlined here:

```{r, eval=F}
vignette("auth", package = "rtweet")
```

- Once you have an account, log in, make an app, give it a name, and copy the following four elements: `api_key`, `api_secret_key`, `access_token`, `access_token_secret`

---

#These Are MINE, Get Your Own!
```{r, eval=F}
api_key <- "abXXRlapnqfhMDhA3n1t00GXaxx"
api_secret_key <- "eyh24B5i8ETIxTGopOL0mQoFnirLnotVEdWeUVTI0eVrXmKwAYxx"
access_token <- "4882273461-PLspkkkZEndhlaJ00uXpQ36KQCK3RefzU61fQACxx"
access_token_secret<-"2ZXjFiwVNEjKPFnWFFxoI3n0Ar0O5Zlio3MyoE1Uc4M4dxx"


## authenticate via web browser
token <- create_token(
  app = "Olgas_scraper",
  consumer_key = api_key,
  consumer_secret = api_secret_key,
  access_token = access_token,
  access_secret = access_token_secret,
  set_renv = TRUE)
```

---

## Friends/Followers

- Twitter [API documentation](https://developer.twitter.com/en/docs/twitter-api/v1/data-dictionary/overview/user-object) distinguishes between *friends* and *followers*

    - *friend* an account a user follows: a limit of 5000
    
    - *follower* account that follows a user: no limit
    

```{r, eval=F}
friends<-get_friends("olga_chyzh", parse=TRUE)
friends

followers<-get_followers("olga_chyzh")
followers

#for accounts with large followings:
flwrs<-get_followers("NateSilver538",n=10000, retryonratelimit = TRUE)
```
    
---

## Tell Me Who Your Friends Are

```{r, eval=F}
olgas_followers<-lookup_users(followers$user_id)
olgas_followers<-select(olgas_followers, user_id, screen_name, description)
```

- Do my friends know each other? 

```{r, eval=F}
lookup_friendships("Pearson_ink","BlaxlandJoel")
```
---
## Make a Network of My First 100 Friends

```{r, eval=F}

olgas_followers<-olgas_followers[1:100,]

#Match each friend with each other friend
friends_net<-expand.grid(screen_name1=olgas_followers$screen_name,screen_name2=olgas_followers$screen_name)

#Add ids and descriptions
friends_net<-friends_net %>% 
  left_join(olgas_followers, by=c("screen_name1"="screen_name")) %>% 
  select(user_id1=user_id, desc1=description, everything()) %>% 
  left_join(olgas_followers, by=c("screen_name2"="screen_name")) %>%
  select(user_id1, user_id2=user_id, screen_name1 , screen_name2, desc2=description, everything()) 


```

---

## Write a Loop to Get Links  
```{r, eval=F}

link<-NULL
for (i in 1:nrow(friends_net)) {
     tryCatch({
  scrnm1<-friends_net[i,3]
  scrnm2<-friends_net[i,4]
  if (scrnm1!=scrnm2){
  fr1<-lookup_friendships(scrnm1,scrnm2)
  link[i]<-ifelse(fr1$value[4]=="TRUE",1,0) #If friend 1 follows friend 2
    } else {link[i]<-NA}
   }, error = function(e) {
      cat("rate limit reached on user: ", i, ", user skipped", fill=TRUE)
     Sys.sleep(901)
     
    }
  )
} 

friends_net$link<-NA
friends_net$link<-link
saveRDS(friends_net,"friends_net.rds")
```

---

## Rate Limits

- Twitter limits how much data you can download at once.

- Use `rate_limit` to find out what the limits are.

```{r, eval=F}
rate_limit(token = NULL, query = "lookup_friendships", parse = TRUE)
```

- Use `Sys.sleep()` to have your loop/program pause once it reaches the rate limit, wait it out and restart.

---
## A Network of My Friends

```{r, eval=TRUE, out.width= "400px"}
friends_net<-readRDS("friends_net.rds")
elist<-friends_net %>% filter(link==1) %>% select(screen_name1,screen_name2)
p<-graph_from_edgelist(as.matrix(elist))
plot(p, vertex.color="turquoise", vertex.size=10, vertex.label=NA)
```

---

## Why Do My Friends Know Each Other?
```{r, eval=T}
olgas_followers<-readRDS("olgas_followers.rds")
olgas_followers <-olgas_followers %>% mutate(description=str_to_lower(description),
                           academ=as.numeric(str_detect(description, "professor")==TRUE |
                                                          str_detect(description, "scientist")==TRUE|
                                                           str_detect(description, "polsci")==TRUE |
                                                           str_detect(description, "polisci")==TRUE |
                                                           str_detect(description, "phd")==TRUE |
                                                           str_detect(description, "political")==TRUE
                                                           ))

```
---

## A Network of My Friends

```{r, eval=F}
mynodes<-as.data.frame(get.vertex.attribute(p)$name)  #get node names in order
names(mynodes)<-"screen_name" 
mynodes %>% left_join(olgas_followers, by="screen_name") %>% mutate(academ=replace_na(academ,0)) -> mynodes #merge in academ
p<-set_vertex_attr(p, "academ", value=mynodes$academ) #set academ as a vertex attribute
V(p)[V(p)$academ == 1]$color <- "purple"
V(p)[V(p)$academ == 0]$color <- "turquoise"
plot(p,  vertex.size=10,  vertex.label=NA, layout=layout.circle)

```
---
## A Network of My Friends

```{r, echo=F}
mynodes<-as.data.frame(get.vertex.attribute(p)$name)  #get node names in order
names(mynodes)<-"screen_name" 
mynodes %>% left_join(olgas_followers, by="screen_name") %>% mutate(academ=replace_na(academ,0)) -> mynodes #merge in academ
p<-set_vertex_attr(p, "academ", value=mynodes$academ) #set academ as a vertex attribute
V(p)[V(p)$academ == 1]$color <- "purple"
V(p)[V(p)$academ == 0]$color <- "turquoise"
plot(p,  vertex.size=10,  vertex.label=NA, layout=layout.circle)

```

---

## Who Are My Non-Academic Friends?

```{r, out.width= "400px"}
mynodes1<- mynodes %>% filter(academ==0) 
elist1<-friends_net %>% filter((screen_name1 %in% mynodes1$screen_name) & (screen_name2 %in% mynodes1$screen_name) & link==1) %>% select(screen_name1,screen_name2)
p<-graph_from_edgelist(as.matrix(elist1))
plot(p,  vertex.size=10, vertex.color="turquoise", vertex.label.dist=2)
```


---
## User Timelines

Specify a user ID or screen name and the number of tweets (max of 3200)

```{r, eval=F}
olgas_tl<-get_timeline("olga_chyzh",n=3200) 
olgas_tl<-olgas_tl %>% 
  mutate(t=ymd_hms(created_at, tz="GMT") , t_ET=with_tz(t,"America/New_York"), hr=hour(t_ET), min=minute(t_ET), time=round((hr*60+min)/5)*5) %>% 
  group_by(time) %>%
  summarise(freq=length(time), .groups="keep") 

olgas_tl %>%  ggplot(aes(x=time,y=freq))+geom_line()+ geom_point()+ scale_x_continuous("",breaks=seq(from=0, to=1380, by=60),labels=seq(from=0,to=23, by=1)) + scale_y_continuous("")+coord_polar() +theme_bw()
table(olgas_tl$freq)
```
---

## When Do I Tweet? When Do I Sleep?

```{r, echo=F}
olgas_tl<-readRDS("olgas_tl.rds")
olgas_tl<-olgas_tl %>% 
  mutate(t=ymd_hms(created_at, tz="GMT") , t_ET=with_tz(t,"America/New_York"), hr=hour(t_ET), min=minute(t_ET), time=round((hr*60+min)/5)*5) %>% 
  group_by(time) %>%
  summarise(freq=length(time), .groups="keep") 

olgas_tl %>%  ggplot(aes(x=time,y=freq))+geom_line()+ geom_point()+ scale_x_continuous("",breaks=seq(from=0, to=1380, by=60),labels=seq(from=0,to=23, by=1)) + scale_y_continuous("")+coord_polar() +theme_bw()

```


---

## A Timeline of a Bot

```{r, eval=F}
bot_tl<-get_timeline("American__Voter",n=3200) 
bot_tl<-bot_tl %>% 
  mutate(t=ymd_hms(created_at, tz="GMT") , t_ET=with_tz(t,"America/New_York"), hr=hour(t_ET), min=minute(t_ET), time=round((hr*60+min)/5)*5) %>% 
  group_by(time) %>%
  summarise(freq=length(time), .groups="keep") 

bot_tl %>%  ggplot(aes(x=time,y=freq))+geom_line()+ geom_point()+ scale_x_continuous("",limits=c(0, 1440),breaks=seq(from=0, to=1380, by=60),labels=seq(from=0,to=23, by=1)) + scale_y_continuous("")+coord_polar() +theme_bw()

table(bot_tl$freq)

```
---

## Does American__Voter Sleep?

```{r, echo=F}
bot_tl<-readRDS("bot_tl.rds")
bot_tl<-bot_tl %>% 
  mutate(t=ymd_hms(created_at, tz="GMT") , t_ET=with_tz(t,"America/New_York"), hr=hour(t_ET), min=minute(t_ET), time=round((hr*60+min)/5)*5) %>% 
  group_by(time) %>%
  summarise(freq=length(time), .groups="keep") 

bot_tl %>%  ggplot(aes(x=time,y=freq))+geom_line()+ geom_point()+ scale_x_continuous("",limits=c(0, 1440),breaks=seq(from=0, to=1380, by=60),labels=seq(from=0,to=23, by=1)) + scale_y_continuous("")+coord_polar() +theme_bw()

```


---

## Tell Me Who Your Friends Are

```{r, eval=F}
library(parallel)
tl_scraper<-function(scrnm){
  tryCatch({
  d<-get_timeline(scrnm,n=1000)
    }, error = function(e) {
      cat(" problem user: ", scrnm, ", user skipped", fill=T)
      })
}

fol_list<-c("andrewheiss","burrelvannjr","CarrieAKJohnson","decustecu","HawkInATX","kristinerudolph","MortazaAbbasi", "shifrinson")
olgas_fol_tl<-do.call(rbind,mclapply(fol_list,tl_scraper))
saveRDS(olgas_fol_tl, "olgas_fol_tl.rds")
```

---
## Tell Me Who Your Friends Are
```{r, eval=F}
olgas_fol_tl %>% 
  mutate(t=ymd_hms(created_at, tz="GMT") , t_ET=with_tz(t,"America/New_York"), hr=hour(t_ET), min=minute(t_ET), time=round((hr*60+min)/10)*10) %>% 
  group_by(time, screen_name) %>%
  summarise(freq=length(time), .groups="keep") %>%  ggplot(aes(x=time,y=freq))+
  geom_line()+ 
  geom_point() + 
  facet_wrap(~screen_name)+ 
  scale_x_continuous("",limits=c(0, 1440),breaks=seq(from=0, to=1380, by=60),labels=seq(from=0,to=23, by=1))+
  scale_y_continuous("")+
  coord_polar() +
  theme_bw()

```
---

## Tell Me Who Your Friends Are
```{r, echo=F, out.width= "400px"}
olgas_fol_tl<-readRDS("olgas_fol_tl3.rds")
olgas_fol_tl %>% 
  mutate(t=ymd_hms(created_at, tz="GMT") , t_ET=with_tz(t,"America/New_York"), hr=hour(t_ET), min=minute(t_ET), time=round((hr*60+min)/10)*10) %>% 
  group_by(time, screen_name) %>%
  summarise(freq=length(time), .groups="keep") %>%  ggplot(aes(x=time,y=freq))+
  geom_line()+ 
  geom_point() + 
  facet_wrap(~screen_name)+ 
  scale_x_continuous("",limits=c(0, 1440),breaks=seq(from=0, to=1380, by=60),labels=seq(from=0,to=23, by=1))+
  scale_y_continuous("")+
  coord_polar() +
  theme_bw()
```

---
## Likes

```{r, eval=F}
olgas_favs <- get_favorites("olga_chyzh", n = 3000) #max is 3000
ts_plot(olgas_favs, "10 days")+theme_bw()
```

```{r, eval=T, echo=F, out.width= "400px"}
olgas_favs <- readRDS("olgas_favs.rds")
ts_plot(olgas_favs, "10 days")+theme_bw()
```


